---
title: "DATA 604: Simulation and Modeling"
author: "Walt Wells, Fall 2017"
subtitle: "Week 6 Discussion"
output:
  html_document:
    css: ../custom.css
    highlight: zenburn
    theme: lumen
---

# Random Variate Generation

## Banks DES 8.8

Develop a generator for a random variable with the pdf below.   Generate 1000 values and plot a histogram. 

\[
f(x)=\begin{cases}
    1/3 & 0 \leq x \leq 2 \\
    1/24 & 2 < x \leq 10 \\
    0 & otherwise \\
      \end{cases}
\]

# Inverse Tranform Method

## Find the CDF 

#### Part 1:  

$$\int_{0}^{x} \frac{1}{3} dx = \frac{x}{3}$$

#### Part 2: 
(Be sure to add in top limit of prior area here) 

$$\int_{2}^{x} \frac{1}{24} dx + \frac{2}{3} = \frac{x}{24} + \frac{7}{12}$$

#### TEST:   
If we insert our top limit 10 do we get back 1?   

$$\frac{10}{24} + \frac{7}{12} = 1$$

#### Therefore our CDF: 

\[
F(x)=\begin{cases}
    0 &  x < 0 \\
    \frac{x}{3} &  0 \leq x \leq 2 \\
    \frac{x}{24} + \frac{7}{12} & 2 < x \leq 10 \\
    1 & x > 10 \\
      \end{cases}
\]

## Set F(X) = R and solve for R

\[
X=\begin{cases}
    3R &  0 \leq R \leq 2/3 \\
    24(R - \frac{7}{12}) &  2/3 < R \leq 1 \\
   \end{cases}
\]

## Generate Random Variate


```{r}
n <- 10 ^ 4
R <- runif(n)
z <- c()

for (i in R) {
   if (i < 2/3) {
       r <- 3 * i
   } else {
       r <- 24 * (i - 7/12)
   }
   z <- c(z, r)
}
```

## Plot

```{r}
hist(z)

```


# References

* [VIDEO - Finding CDF from PDF](https://www.youtube.com/watch?v=WDGyFeuk9oI)
