---
title: "DATA 604: Simulation and Modeling"
author: "Walt Wells, Fall 2017"
subtitle: "Week 5 Discussion"
output:
  html_document:
    css: ../custom.css
    highlight: zenburn
    theme: lumen
---

# Random Number Generation

This week, one of the topics in random number generation. Using the software / spreadsheet of your choice, build one of the random number generators discussed in Banks DES. Post your code to the discussion board. What challenges did you have?

# Environment Prep

```{r, message=F, warning=F}
if (!require('stats')) install.packages('stats')
if (!require('ggplot2')) install.packages('ggplot2')
if (!require('gridExtra')) install.packages('gridExtra')
```

# Generate Random Number Using Linear Congruential Method

```{r}
lcm <- function(seed, a, c, m, n) {
    # Use Linear Congruential Method to generate Random Numbers
    #
    # Args: 
    #   seed:  set X_0 or initial value
    #   a: multiplier
    #   c: increment
    #   m: modulus
    #   n: number random numbers desired
    # 
    # Returns:
    #   list of length n with random numbers
    
    list <- c()
    
    Xi <- seed
    
    for(i in 0:n) {
        Xi <- (a*Xi + c) %% m
        Ri <- Xi / m
        list[i+1] <- Ri
    }
    
    list
}
```


## Run LCM method

### Bad Random 

We'll try and generate a string of random variables but with a short period so that they are likely to repeat.

```{r}
X_0 <- 2
a <- 8
c <- 2
m <- 100
BadRand <- lcm(X_0, a, c, m, 10^5-1)
```

```{r}
a <- ggplot() + geom_density(aes(BadRand))
b <- ggplot() + stat_ecdf(aes(BadRand))
grid.arrange(a, b, ncol=2)
```

### Good Random

We will use the work of Learmonth and Lewis, 1973; and Lewis et al., 1969 to set our variables to ensure very long period. 

```{r}
X_0 <- 123457
a <- 7^5
c <- 0
m <- 2^31 - 1
GoodRand <- lcm(X_0, a, c, m, 10^5-1)
```

```{r}
a <- ggplot() + geom_density(aes(GoodRand))
b <- ggplot() + stat_ecdf(aes(GoodRand))
grid.arrange(a, b, ncol=2)
```

## Did we generate random numbers that maintain the appearance of randomness?

### Simple Frequency Check

```{r}
CountReps <- function(Rand) {
    # checks if any randomly generated numbers appeared more than once
    #
    # Args:  
    #   Rand: string of random numbers to test
    #
    # Results:
    #   returns df, if there are 0 rows, there are zero repeats
    
    t <- data.frame(table(Rand))
    t <- t[t$Freq > 1,]
    t
}
```

```{r}
CountReps(GoodRand)
CountReps(BadRand)
```

