---
title: "Simulation and Modeling"
author: "Walt Wells, Fall 2017"
subtitle: "Input Analysis / Variance Reduction Pt1"
output:
  html_document:
    css: ../custom.css
    highlight: zenburn
    theme: lumen
---

# Environment Prep

```{r, message=F, warning=F}
if (!require('xlsx')) install.packages('xlsx')
if (!require('fitdistrplus')) install.packages('fitdistrplus')
if (!require('dplyr')) install.packages('dplyr')
if (!require('ggplot2')) install.packages('ggplot2')
if (!require('gridExtra')) install.packages('gridExtra')
```

# Helper Function - Fitting/Plotting

We'll use this a few times to test fit using the fitdistrplus packge, so let's make a function. 

```{r}
GoodnessPlot <- function(vector, try_list, discrete=FALSE) {
    # given a data distribution uses the fitdistrplus package to 
    # plot goodness of fits agains a list of distributions to try
    #
    # Args: 
    #   vector:  desired distribution
    #   try_list:  list of distributions to test
    #
    # Returns:
    #   a series of plots showing:
    #     Density - closeness of fit for each distribution
    #     QQ - how do the tails match
    #     CDF - how do the cdfs match
    #     PP - how well does the center match
    #   a series of goodness of fit tests
    #      Cramer-von Mises, Kolmogorov-Smirnov and Anderson-Darling

    plot.legend <- try_list
    n <- length(try_list)
    fit_list <- list()
    
    for (i in 1:length(try_list)) {
        name <- paste0("fit", i)
        if (discrete) {
            fit <- fitdist(vector, try_list[i], discrete=T) 
            } else {
            fit <- fitdist(vector, try_list[i])
            }
        fit_list[[name]] <- fit
    }
    
    denscomp(fit_list, legendtext = plot.legend)
    qqcomp(fit_list, legendtext = plot.legend)
    cdfcomp(fit_list, legendtext = plot.legend)
    ppcomp(fit_list, legendtext = plot.legend)
    
    gofstat(fit_list)
}

```


# Kelton 6.5.1

```{r}
df1 <- read.xlsx('Problem_Dataset_06_01.xls', 1, header=F)
plotdist(df1$X1, histo = TRUE, demp = TRUE)
descdist(df1$X1, boot=500)

try <- c("gamma", "weibull", "lnorm", "logis")
GoodnessPlot(df1$X1, try)
```

It looks like the best option here is the Lognormal Distribution.  

```{r}
fln <- fitdist(df1$X1, "lnorm")
coef <- round(coef(fln), 4)
RV <- rlnorm(10^5, coef[1], coef[2])

hist(df1$X1, prob=TRUE, col="grey", ylim=c(0, .09), main="Density of Data vs Fit of Chosen Distribution")
lines(density(RV), col="blue", lwd=2) 

print(paste0("Simio Command:  Random.Lognormal(", coef[1], ", ", coef[2], ")"))
```


# Kelton 6.5.2

```{r}
df2 <- read.xlsx('Problem_Dataset_06_02.xls', 1, header=F)
plotdist(df2$X1, histo = TRUE, demp = TRUE)
descdist(df2$X1, boot=500)

try <- c("gamma", "weibull", "lnorm", "logis")
GoodnessPlot(df2$X1, try)
```

It looks like the best option here is again the Lognormal Distribution.  

```{r}
fln <- fitdist(df2$X1, "lnorm")
coef <- round(coef(fln), 4)
RV <- rlnorm(10^5, coef[1], coef[2])

hist(df2$X1, prob=TRUE, col="grey", ylim=c(0, .11), main="Density of Data vs Fit of Chosen Distribution")
lines(density(RV), col="blue", lwd=2) 

print(paste0("Simio Command:  Random.Lognormal(", coef[1], ", ", coef[2], ")"))
```

# Kelton 6.5.3

```{r}
df3 <- read.xlsx('Problem_Dataset_06_03.xls', 1, header=F)
plotdist(df3$X1, histo = TRUE, demp = TRUE)
descdist(df3$X1, boot=500)

try <- c("gamma", "weibull", "lnorm")
GoodnessPlot(df3$X1, try, discrete=TRUE)
```

Here let's try the Gamma Distribution.

```{r}
fln <- fitdist(df3$X1, "gamma", discrete=TRUE)
coef <- round(coef(fln), 4)
RV <- rgamma(10^5, coef[1], coef[2])

hist(df3$X1, prob=TRUE, col="grey", ylim=c(0, .44), main="Density of Data vs Fit of Chosen Distribution")
lines(density(RV), col="blue", lwd=2) 

print(paste0("Simio Command:  Random.Gamma(", coef[1], ", ", coef[2], ")"))
```

# Kelton 6.5.4



# Kelton 6.5.5


# Kelton 6.5.8



# References

* https://cran.csiro.au/web/packages/fitdistrplus/vignettes/paper2JSS.pdf