---
title: "Simulation and Modeling"
author: "Walt Wells, Fall 2017"
subtitle: "Input Analysis / Variance Reduction Pt2"
output:
  html_document:
    css: ../custom.css
    highlight: zenburn
    theme: lumen
---

# Environment Prep

```{r, message=F, warning=F}
if (!require('dplyr')) install.packages('dplyr')
if (!require('ggplot2')) install.packages('ggplot2')
if (!require('gridExtra')) install.packages('gridExtra')
```

# Kelton 6.5.4

### CDF for Continuous Uniform Distribution

\[
F(x)=\begin{cases}
    0 &  x < a \\
    \frac{x-a}{b-a} &  a \leq x \leq b \\
    1 & x > b \\
      \end{cases}
\]

### Inverse Transform (Set F(x) = R, solve for X)

\[
X= a + (b-a)R 
\]

# Kelton 6.5.5

### CDF for Weibull Distribution

\[
F(x)=\begin{cases}
    1 - e^{-(x/\lambda)^{k}}&  x \geq 0 \\
    0 & x < 0 \\
      \end{cases}
\]

### Inverse Transform (Set F(x) = R, solve for X)

\[
X= \lambda[-ln(1-R)]^{1/k}
\]

# Kelton 6.5.8 Revisited

First let's organize what we know. 

```{r}
buy <- c(1.05, 3.17, 1.99, 0.95) #dollars
sell <- c(1.29, 3.76, 2.23, 1.65) #dollars

df <- data.frame(cbind(buy, sell))
rownames(df) <- c("oats", "peas", "beans", "barley")

oatDist <- data.frame(cbind(
    lbs = c(0, 0.5, 1, 1.5, 2, 3, 4, 5, 7.5, 10),
    prob = c(0.05, 0.07, 0.09, 0.11, 0.15, 0.25, 0.1, 0.09, 0.06, 0.03)
))

peaDist <- data.frame(cbind(
    lbs = c(0, 0.5, 1, 1.5, 2, 3),
    prob = c(0.1, 0.2, 0.2, 0.3, 0.1, 0.1)
))

beansDist <- data.frame(cbind(
    lbs = c(0, 1, 3, 4.5),
    prob = c(0.2, 0.4, 0.3, 0.1)
))

barleyDist <- data.frame(cbind(
    lbs = c(0, 0.5, 1, 3.5),
    prob = c(0.2, 0.4, 0.3, 0.1)
))
```

### Create CumProbability

```{r}
oatDist$cumProb <- cumsum(oatDist$prob)
peaDist$cumProb <- cumsum(peaDist$prob)
beansDist$cumProb <- cumsum(beansDist$prob)
barleyDist$cumProb <- cumsum(barleyDist$prob)
```


## 90 Day Simulation, Each Product

Now we build a function that will run our simulation.  This function has two helpers that a)  generate a random sample from a discrete uniform distribution, then generates the appropriate demand based on the CDF, b) gets the profit, revenue, and cost for a given day.


```{r}
simMarket <- function(duration, crop, cropDist) {
    # return a time series dataframe where each 
    # row describes results for crop given a random demand
    #
    # Args: 
    #   duration: length of simulation
    #   crop:  string representing crop type
    #   cropDist:   probability of a crop's demand
    #
    # Returns:
    #   results:  df showing costs, revenue, profit, day number, and demand
    
    # Helper Function:  Generate Random Variable
    getDemand <- function(cropDist) {
        # get a sample from a discrete rand distribution representing 
        # demand for a given crop
        #
        # Args: 
        #   cropDist:  probable distribution for crop type
        #
        # Returns:
        #   demand:  number representing random day's demand
        
        R <- runif(1)
        demand <- cropDist[which(cropDist[,3] >= R), ][1,1]
        demand
    }
    
    # Helper Function:  Get Cost, Revenue and Profit, Single Day
    getDay <- function(demand, crop) {
        # return a df row showing results for a day's sales
        #
        # Args: 
        #   demand:  that day's demand
        #   crop:  string representing crop type
        #
        # Returns:
        #   day:  df row showing day's cost, revenue, profit 
        #         for a crop given demand
        
        row <- df[crop, ]
        day <- data.frame(cost = rep(0,1), 
                          revenue = rep(0,1),
                          profit = rep(0,1))
        day$cost <- -(row$buy * demand)
        day$revenue <- row$sell * demand
        day$profit <- day$revenue + day$cost
        day
    }
    
    # run helpers to get final results
    results <- data.frame()
    for(i in 1:duration) {
        dayDemand <- getDemand(cropDist)
        row <- getDay(dayDemand, crop)
        row$demand <- dayDemand
        row$day <- i
        results <- rbind(results, row)
    }
    results
}
```

## Run Sim for each crop

```{r}
oats <- simMarket(90, 'oats', oatDist)
peas <- simMarket(90, 'peas', peaDist)
beans <- simMarket(90, 'beans', beansDist)
barley <- simMarket(90, 'barley', barleyDist)

# example 
head(oats, 10)
```

## Helper Function:  Make Plots

```{r}
getPlot <- function(df) {
    # return a line plot showing sim results over time frame
    #
    # Args: 
    #   df: data frame with the 90 day sim results for a given crop
    #
    # Returns:
    #   p:  ggplot object
    
    profit <- sum(df$profit)
    crop <- deparse(substitute(df))
    substr(crop, 1, 1) <- toupper(substr(crop, 1, 1))
    title <- paste0("Sim Results: ", crop)
    subtitle <- paste0("Total Profit: $", profit)
    
    p <- ggplot(df, aes(x=day)) + 
            geom_line(aes(y=revenue, col="revenue")) + 
            geom_line(aes(y=cost, col="cost")) + 
            geom_line(aes(y=profit, col="profit")) + 
            labs(title=title,
                 subtitle=subtitle, 
                 y="dollars",
                 x="days")
    p
}
```

## Visualize One Simulation

Walther got a lot poorer. 

```{r}
oatPlot <- getPlot(oats)
peasPlot <- getPlot(peas)
beansPlot <- getPlot(beans)
barleyPlot <- getPlot(barley)
grid.arrange(oatPlot, peasPlot, beansPlot, barleyPlot, ncol=2)
```

## Repeat Sim

```{r}
n <- 100

oatList <- t(replicate(n, simMarket(90, 'oats', oatDist)))
oatSimMean <- mean(unlist(oatList[,3])) * 90
peasList <- t(replicate(n, simMarket(90, 'peas', peaDist)))
peasSimMean <- mean(unlist(peasList[,3])) * 90
beansList <- t(replicate(n, simMarket(90, 'beans', beansDist)))
beansSimMean <- mean(unlist(beansList[,3])) * 90
barleyList <- t(replicate(n, simMarket(90, 'barley', barleyDist)))
barleySimMean <- mean(unlist(barleyList[,3])) * 90
```

## Means of Profits

```{r}
profitMeans <- round(data.frame(oats = oatSimMean,
                          peas = peasSimMean,
                          beans = beansSimMean,
                          barley =  barleySimMean), 2)

## Means of Profits Across all Rep
knitr::kable(profitMeans)
```


# References

* Banks, et all: Discrete Event Simulation