---
title: "DATA 604: Simulation and Modeling"
author: "Walt Wells, Fall 2017"
subtitle: "Week 9 Discussion"
output:
  html_document:
    css: ../custom.css
    highlight: zenburn
    theme: lumen
---

# Monte Carlo Integration and Variance Reduction

## 5.13

Find two importance functions f1 and f2 that are supported on (1, $\infty$) and are close to

$$g(x) = \frac{x^2}{\sqrt{2\pi}} e^{-x^2/2}, x>1$$

Which of the two should produce a smaller variation in estimating: 

$$g(X) = \int_{1}^{\infty} \frac{x^2}{\sqrt{2\pi}} e^{-x^2/2} dx$$

```{r}
g <- function(x) {
    (x^2/sqrt(2*pi)) * exp(-(x^2)/2) 
    } 
```


Here's a look at the function and relative area of interest.   

```{r}
par(mfrow=c(1,2))
curve(g, -5, 5, main="Unbounded g(x)")

x <- seq(1,5, 0.01)
plot(g(x)~x, type="h", main="Bounded g(x)")
par(mfrow=c(1,1))
```

Let's see what might be some good options for importance functions

```{r}
f1 <- function(x) {
    exp(-(x^2)/2) 
    } 
f2 <- function(x) {
    (1/sqrt(2*pi)) * exp(-(x^2)/2)
}
```

```{r}
x <- seq(1,10, 0.01)
plot(x, f1(x)/g(x), type="l", main="", ylab="fi(x)/g(x)", lty=1, ylim=c(0, 1))
lines(x, f2(x)/g(x), lty=2)
legend("topright", legend=c("f1", "f2"), lty=1:2) 
```


# References

* Statistical Computing with R, Maria L. Rizzo, Chapter 5
* Judd Andermann:  Week 8 http://rpubs.com/janderman/data604_wk8_discuss
* https://math.stackexchange.com/questions/1200443/evaluating-difficult-monte-carlo-integration-in-r
