---
title: "DATA 604: Simulation and Modeling"
author: "Walt Wells, Fall 2017"
subtitle: "Week 3 Discussion"
output:
  html_document:
    css: ../custom.css
    highlight: zenburn
    theme: lumen
---

# Environment Prep

```{r, message=F, warning=F}
if (!require('dplyr')) install.packages('dplyr')
if (!require('ggplot2')) install.packages('ggplot2')
if (!require('gridExtra')) install.packages('gridExtra')
```

# 3.5.17 Walther's Produce Stand

First let's organize what we know. 

```{r}
buy <- c(1.05, 3.17, 1.99, 0.95) #dollars
sell <- c(1.29, 3.76, 2.23, 1.65) #dollars
demandTop <- c(10, 8, 14, 11) #lbs

df <- data.frame(cbind(buy, sell, demandTop))
rownames(df) <- c("oats", "peas", "beans", "barley")
df
```

## 90 Day Simulation, Each Product

Now we build a function that will run our simulation.  This function has two helpers that a)  generate a random sample from a discrete uniform distribution representing daily demand, b) gets the profit, revenue, and cost for a given day.


```{r}
simMarket <- function(duration, crop) {
    # return a time series dataframe where each 
    # row describes results for crop given a random demand
    #
    # Args: 
    #   duration: length of simulation
    #   crop:  string representing crop type
    #
    # Returns:
    #   results:  df showing costs, revenue, profit, day number, and demand
    
    # Helper Function:  Generate Random Variable
    getDemand <- function(crop) {
        # get a sample from a discrete rand distribution representing 
        # demand for a given crop
        #
        # Args: 
        #   crop:  string representing crop type
        #
        # Returns:
        #   demand:  number representing random day's demand
        
        row <- df[crop,]
        rangeTop <- row$demandTop
        demand <- ceiling(runif(1, -1, rangeTop)) #-1 so we include a demand of 0
        demand
    }
    
    # Helper Function:  Get Cost, Revenue and Profit, Single Day
    getDay <- function(demand, crop) {
        # return a df row showing results for a day's sales
        #
        # Args: 
        #   demand:  that day's demand
        #   crop:  string representing crop type
        #
        # Returns:
        #   day:  df row showing day's cost, revenue, profit 
        #         for a crop given demand
        
        row <- df[crop, ]
        day <- data.frame(cost = rep(0,1), 
                          revenue = rep(0,1),
                          profit = rep(0,1))
        day$cost <- -(row$buy * demand)
        day$revenue <- row$sell * demand
        day$profit <- day$revenue + day$cost
        day
    }
    
    # run helpers to get final results
    results <- data.frame()
    for(i in 1:duration) {
        dayDemand <- getDemand(crop)
        row <- getDay(dayDemand, crop)
        row$demand <- dayDemand
        row$day <- i
        results <- rbind(results, row)
    }
    results
}
```

## Run Sim for each crop

```{r}
oats <- simMarket(90, 'oats')
peas <- simMarket(90, 'peas')
beans <- simMarket(90, 'beans')
barley <- simMarket(90, 'barley')

# example 
head(oats, 10)
```


## Helper Function:  Make Plots

```{r}
getPlot <- function(df) {
    # return a line plot showing sim results over time frame
    #
    # Args: 
    #   df: data frame with the 90 day sim results for a given crop
    #
    # Returns:
    #   p:  ggplot object
    
    profit <- sum(df$profit)
    crop <- deparse(substitute(df))
    substr(crop, 1, 1) <- toupper(substr(crop, 1, 1))
    title <- paste0("Sim Results: ", crop)
    subtitle <- paste0("Total Profit: $", profit)
    
    p <- ggplot(df, aes(x=day)) + 
            geom_line(aes(y=revenue, col="revenue")) + 
            geom_line(aes(y=cost, col="cost")) + 
            geom_line(aes(y=profit, col="profit")) + 
            labs(title=title,
                 subtitle=subtitle, 
                 y="dollars",
                 x="days")
    p
}
```

## Visualize simulation

```{r}
oatPlot <- getPlot(oats)
peasPlot <- getPlot(peas)
beansPlot <- getPlot(beans)
barleyPlot <- getPlot(barley)
grid.arrange(oatPlot, peasPlot, beansPlot, barleyPlot, ncol=2)
```

## What are the totals for the length of our sim? 

```{r}
all <- rbind(oats, peas, beans, barley)
dailySum <- all %>%
    group_by(day) %>%
    summarise(cost = sum(cost),
              revenue = sum(revenue),
              profit = sum(profit))
colSums(dailySum[2:4])
```
