---
title: "DATA 604: Simulation and Modeling"
author: "Walt Wells, Fall 2017"
subtitle: "Week 10 Discussion"
output:
  html_document:
    css: ../custom.css
    highlight: zenburn
    theme: lumen
---

# Markov Chain Monte Carlo - Metropolis Sampler, Laplace Distribution

Implement a random walk Metropolis sampler for generating the standard Laplace distribution. For the increment, simulate from a normal distribution. Compare the chains generated when different variances are used for the proposal distribution. Also, compute the acceptance rates of each chain.

### Standard Laplace Distribution: 

$$g(x) = \frac{1}{2}e^{-|x|}, x \in {R}$$

## Environment Prep

```{r, warning=F, message=F}
if (!require('rmutil')) install.packages('rmutil') #for dplaplace
```

### Helper Function

```{r}
rw.Metropolis <- function(sigma, x0, N) {
    x <- numeric(N)
    x[1] <- x0
    u <- runif(N)
    k <- 0
    for (i in 2:N) {
        y <- rnorm(1, x[i-1], sigma)
        if (u[i] <= dlaplace(y) / dlaplace(x[i-1])) {
            x[i] <- y 
        } else {
            x[i] <- x[i-1]
            k <- k+1
        }
    }
    return(list(x=x, k=k))
}
```

### Run Metropolis Sampler MCMC

```{r}
set.seed(604)
N <- 5000
sigma <- c(.05, .5, 1, 5)

x0 <- 25
rw1 <- rw.Metropolis(sigma[1], x0, N)
rw2 <- rw.Metropolis(sigma[2], x0, N)
rw3 <- rw.Metropolis(sigma[3], x0, N)
rw4 <- rw.Metropolis(sigma[4], x0, N)
```

### Calculate Acceptance Rates

```{r}
#num rejected points
rejected <- c(rw1$k, rw2$k, rw3$k, rw4$k)

#show % accepted
perRej <- data.frame(lapply(rejected, function(x) round((1-(x/N)) * 100, 2)))
names(perRej) <- c("Sigma.05", "Sigma.5", "Sigma1", "Sigma5")
knitr::kable(perRej)
```

## Plot Each Chain

```{r}
par(mfrow=c(2,2))
refline <- qt(c(.025, .975), df=2)
rw <- cbind(rw1$x, rw2$x, rw3$x, rw4$x)
for (j in 1:4) {
    plot(rw[, j], type='l', 
        xlab=bquote(sigma == .(round(sigma[j], 3))),
        ylab = "X", ylim=range(rw[,j]))
    abline(h=refline)
}
```

### Compare Values at different N in chain

```{r}
a <- c(.05, seq(.1, .9, .1), .95)
Q <- qlaplace(a)
rw <- cbind(rw1$x, rw2$x, rw3$x, rw4$x)
Qrw <- apply(rw, 2, function(x) quantile(x, a))
chains <- data.frame(round(cbind(Q, Qrw), 3))
names(chains) <- c("Laplace/Target", "Sigma.05", "Sigma.5", "Sigma1", "Sigma5")
knitr::kable(chains)
```

# References

* Statistical Computing with R, Maria L. Rizzo, Chapter 5

